---
title: "Enzyme Kinetics Lab"
author: "Nathan Malmberg"
date: "6/17/2019"
output:
  bookdown::html_document2:
    number_sections: FALSE
bibliography: [knitr.bib, references.bib]
csl: american-chemical-society.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(units)
units_options(group=c("(",")"))
```

## Week 1

Use the following code block to keep track of the masses, volumes, etc. of the
solutions you create for lab. Because you will have to make many of these solutions
both weeks of the lab, it will be convenient to keep track of the reagents in vectors,
so the math only needs to be done once.

```{r reagents}
### This lab will use units extensively, so it will be worth learning
### how to manage numbers with units using the 'units' package in R.
### The most important aspect is associating numbers with units. This
### can be done in multiple ways, two of which will be used here. One
### is to assign a unit string to the units function, like:
### units(soda) <- "oz"
### The other is the set_units function, which has two arguments: the
### number and the units for the number. To demonstrate this, the
### reaction volume will be created with the set_units function. This
### volume should be the same for all the reactions for this experiment.
ReactionVolume <- set_units(3,mL)

### Stock Solutions
### The buffer can be made from a concentrated stock, but fresh solutions
### of enzyme, L-Dopa, D-Dopa, cinnamic acid and thiourea need to be made
### each week. Record the
### masses of reagent and volumes of buffer used to make these solutions
### here, so you can calculate the concentrations of the reagents in the
### stock solutions. The L-Dopa solution needs to be made both weeks, but
### the D-Dopa, cinnamic acid and thiourea solutions should only need to
### be made once.
LDopaMass <- set_units(c(,),g)
LDopaVolume <- set_units(c(,),mL)
DDopaMass <- set_units(,g)
DDopaVolume <- set_units(,mL)
cinnamicMass <- ;
cinnamicVolume <- ;
thioureaMass <- ;
thioureaVolume <- ;
### Calculate the molar concentrations of the stock solutions using the
### mass and molecular weights of the reagents, and the volumes of the
### solutions.
DopaFW <- set_units(,g/mol)
LDopaStock <- ;
DDopaStock <- ;
cinnamicFW <- ;
thioureaFW <- ;
cinnamicStock <- ;
thioureaStock <- ;
```

Parts A and B need to be done each week of the lab, but we'll include it in Week 1
for the purposes of this template document.

```{r PartsAB}
### Part A: Measure Enzyme Concentration
### Since we do this part each week, we can store the absorbances
### as a length two vector, and do the calculations all at once.
EnzymeAbsorbance <- set_units(c(,))
### Calculate molar enzyme concentration using the absorbance,
### the extinction coefficient for the enzyme, and the molecular
### weight of the enzyme. This is a good exercise for unit conversions.
EnzymeE <- set_units(,L/g)
EnzymeFW <- set_units(,g/mol)
EnzymeStock <- ;

### Part B: Measure Optimal Enzyme Amount
### When you determine the volume of enzyme needed for each reaction for
### parts C and D, record the values here. There should be two values in
### the vector, one for each week.
EnzymeVolumes <- set_units(c(,),uL)
### Because the concentration of enzyme in the stock solution is not the
### same as the concentration of the enzyme in the reaction, you'll need
### to calculate the concentration of enzyme in the reaction.
EnzymeReaction <- ;
```

Part C is the main part of the first week's lab. We'll perform calculations in one
chunk, and make a plot in another. Note that most of these calculations are not
complete, although in some cases the most technically difficult part has been
supplied.

```{r substratecalc, results="hide"}
### Part C: Comparison of Substrates
### For this part of the experiment, record the identity of the substrate,
### the volume of substrate added to each reaction, and the slope of the
### best-fit line for the absorbance measured as a function of time (s).
### If you use the enzymekinetics.py program on the Genesys machines, this
### will be saved as a CSV file that you can copy into the function below.
### If you separate values with spaces rather than commas, change "csv"
### to "table".
Substrates <- read.csv(header=TRUE,text="")
### Assign appropriate units to each of the columns
units(Substrates$Substrate.Vol) <- "mL"
units(Substrates$Slope) <- ;
### The substrate concentration in the reaction is based on the concentration
### of the stock solution as well as the volume of substrate added. Be sure
### to use the concentration of stock solution from the correct week.
Substrates$Concentration <- LDopaStock[1] #Incomplete
### The following line can be used to change the Concentration values for
### those rows that use D-Dopa as a substrate, if the concentration of the two
### stock solutions is different.
Substrates$Concentration[Substrates$Substrate=="D-Dopa"] = DDopaStock *
    Substrates$Substrate.Vol[Substrates$Substrate=="D-Dopa"] #Incomplete
### It will be helpful to specify the absorptivity of the product to aid
### your calculations. Be sure to supply appropriate units.
DopachromeE <- ;
### The line below presumes that the column of data with the slopes of the lines
### of best fit for each reaction is called Slope. Adjust if you've called it
### something else.
Substrates$Rate <- Substrates$Slope #Incomplete
### Fit the Rate and Concentration to the Michaelis-Menton equation using
### nonlinear least-squares curve fitting, once for each substrate. Save the fits
### in objects that can be used for extracting parameters or plotting curves.
Ldopafit <- nls(Rate ~ SSmicmen(Concentration,Vmax,Km), data=drop_units(Substrates),
                subset=Substrate=="L-Dopa")
Ddopafit <- nls(Rate ~ SSmicmen(Concentration,Vmax,Km), data=drop_units(Substrates),
                subset=Substrate=="D-Dopa")
### View the parameters to compare Vmax and Km.
summary(Ldopafit)
summary(Ddopafit)
### We can calculate turnover number for each curve, since we know the enzyme
### concentration, as well.
LdopaVmax <- coef(Ldopafit)[["Vmax"]]
units(LdopaVmax) <- deparse_unit(Substrates$Rate)
DdopaVmax <- coef(Ddopafit)[["Vmax"]]
units(DdopaVmax) <- deparse_unit(Substrates$Rate)
Ldopakcat <- ;
Ddopakcat <- ;
### And since we have Km as well, we can calculate the specificity constant.
LdopaKm <- coef(Ldopafit)[["Km"]]
units(LdopaKm) <- deparse_unit(Substrates$Concentration)
DdopaKm <- coef(Ddopafit)[["Km"]]
units(DdopaKm) <- deparse_unit(Substrates$Concentration)
LdopaSC <- ;
DdopaSC <- ;
```

It will make the most sense to plot both curves on one plot, to make comparison
easier. This plot will require a detailed caption, which can be written below.

(ref:substrate) Write your more detailed caption here.

```{r subplot, fig.cap='(ref:substrate)', fig.asp=0.7}
### Create a plot of the reaction rate as a function of substrate concentration.
### To make sure all the points fit on the plot, we'll create a blank plot
### so that the axes are scaled correctly, and then add separate points with
### different symbols.
plot(Rate ~ Concentration, data=Substrates, type="n",
     xlab=make_unit_label("'[Dopa]'",Substrates$Concentration),
     ylab=make_unit_label("V[0]",Substrates$Rate))
points(Rate ~ Concentration, data=Substrates, pch=1, subset=Substrate=="L-Dopa")
points(Rate ~ Concentration, data=Substrates, pch=2, subset=Substrate=="D-Dopa")
### Draw the lines for the curve fit for each substrate.
curve(predict(Ldopafit, newdata=list(Concentration=x)), lty=1, add=TRUE)
curve(predict(Ddopafit, newdata=list(Concentration=x)), lty=2, add=TRUE)
### You'll need to adjust the x and y values for the legend, so it's visible.
legend(x="topleft",legend=c("L-Dopa","D-Dopa"),pch=1:2,lty=1:2)
```

## Week Two

Remember to complete parts A and B again, though you can put your data into the
PartsAB chunk above.

As in Week One, we'll divide up the analysis of data into calculations and
plot. This week will also be helped by the use of some functions written
by Dr. Malmberg to analyze enzyme kinetics in the presence of an inhibitor.
These functions are found in a package that may be installed from [GitHub](http://github.com/nmalmberg/SSenzymekinetics), or from a package file
downloaded from Canvas. The latter option can be done from the Tools menu of
RStudio.

```{r inhcalc, results="hide"}
library(SSenzymekinetics)
inhibitordata <- read.csv(header=TRUE, text="")

### Set units for each column that has units
units(inhibitordata$Substrate) <- ;
units(inhibitordata$Inhibitor) <- ;
units(inhibitordata$Slope) <- ;

### Calculate substrate concentration using Week 2 stock concentration
### of L-Dopa and the volume added to the reaction.
inhibitordata$SubCon <- inhibitordata$Substrate # Incomplete
inhibitordata$InhCon <- inhibitordata$Inhibitor # Incomplete
### Calculate the reaction rate based on the slopes of the lines.
inhibitordata$Rate <- inhibitordata$Slope # Incomplete
### Model the reaction as though there were no effect by the inhibitor.
noinhibitnls <- nls(Rate~SSmicmen(SubCon, Vmax, Km),
                    data=drop_units(inhibitordata))
### Fit the data to competitive, uncompetitive and mixed inhibition models.
compnls <- nls(Rate~SScompinh(SubCon, InhCon, Vmax, Km, Kc),
               data=drop_units(inhibitordata))
unnls <- nls(Rate~SSuncompinh(SubCon, InhCon, Vmax, Km, Ku),
             data=drop_units(inhibitordata))
mixednls <- nls(Rate~SSmixedinh(SubCon, InhCon, Vmax, Km, Kc, Ku),
                data=drop_units(inhibitordata))
### Compare the residuals from the competitive and uncompetitive inhibition.
### The fit with smaller residuals is the better fit.
summary(compnls)
summary(unnls)
### Perform an analysis of variance to see whether the model is significantly
### improved by including inhibition at all, or by including both types of
### inhibition. Replace compnls with unnls if uncompetitive was a better fit.
inhanalysis <- anova(noinhibitnls, compnls, mixednls)
inhanalysis
```

It might be helpful to include a table with the results of the analysis of
variance, for the purposes of discussing your results.

```{r}
knitr::kable(inhanalysis, caption = "(ref:analysis)")
```

(ref:analysis) Include your descriptive table caption for the anova here.

```{r inhibitorplot, fig.cap="(ref:inhibitor)",fig.asp=0.7}
plot(Rate ~ SubCon, data=inhibitordata, type="n",
     xlab=make_unit_label("'[Dopa]'",inhibitordata$SubCon),
     ylab=make_unit_label("V[0]",inhibitordata$Rate))
points(Rate ~ SubCon, data=inhibitordata, pch=1,
       subset=(Inhibitor==set_units(0,mL)))
### You may have to adjust Inhibitor if you used a different volume.
points(Rate ~ SubCon, data=inhibitordata, pch=2,
       subset=(Inhibitor==set_units(1,mL)))
### Draw curves for the different models you've fit, both for the points
### without inhibitor and with inhibitor. Change compnls to unnls if you
### decided that fit better.
curve(predict(noinhibitnls, newdata=list(SubCon=x)), lty=1, add=TRUE)
curve(predict(compnls, newdata=list(SubCon=x, InhCon=0)), lty=2, add=TRUE)
curve(predict(compnls, newdata=list(SubCon=x,
                                    InhCon=max(drop_units(inhibitordata$InhCon)))),
      lty=2, add=TRUE)
curve(predict(mixednls, newdata=list(SubCon=x, InhCon=0)), lty=3, add=TRUE)
curve(predict(mixednls, newdata=list(SubCon=x,
                                     InhCon=max(drop_units(inhibitordata$InhCon)))),
      lty=3, add=TRUE)
### You'll need to adjust the x and y values for the legend, so it's visible.
legend(x="topleft",legend=c("-Phe","+Phe","None", "Competitive", "Mixed"),
       pch=c(1:2,NA,NA,NA),lty=c(0,0,1:3))
```

Of course, you will also need a caption for the inhibition plot.

(ref:inhibitor) Put your inhibition plot caption here.

Of course, you also should have some references in this lab report.

```{r bibliography, include=FALSE}
knitr::write_bib(c("base","units","SSenzymekinetics"), "knitr.bib")
```

## References
