---
title: "Lipid Lab"
author: "Nathan Malmberg"
date: "6/28/2019"
output: 
  bookdown::html_document2:
    number_sections: FALSE
bibliography: [knitr.bib, references.bib]
csl: american-chemical-society.csl
abstract: (ref:abstract)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

(ref:abstract) Replace this sentence with your abstract.


This lab report will require an introduction that has at least two citations
of the primary literature, so be sure to add those to the references.bib file.

Use the code block below for your calculations of the relative mobilities of
the different lipids run on the TLC plate. The block will create a table
of mobility factors for each lipid.

(ref:tlctable) Write your caption for your TLC table here.

```{r tlc}
tlcdata <- read.csv(header = TRUE, stringsAsFactors = TRUE, text = "
Lipid,Distance")
# You'll need to calculate the mobility factors before the table can print them

knitr::kable(tlcdata[c("Lipid","Rf")],caption = "(ref:tlctable)")
```

The code block below will create a plot of your TLC trace. Make sure you save
your trace file to your project folder.

```{r gc, fig.cap="(ref:gc)"}
gctable <- read.table("nutmeg.xy", header = FALSE, col.names = c("Time","Response"))
plot(Response~Time, data = gctable, xlab="Time (min)", ylab = "Response",
     pty=0, lty=1)
```

(ref:gc) Write your caption for the GC trace here.

As always, don't forget to cite your sources as appropriate.
Make sure you have two articles from the primary literature in your
references file to cite in your introduction.

```{r references, include=FALSE}
knitr::write_bib(c("base"), "knitr.bib")
```

## References

<div id="refs"></div>

## Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```

