---
title: "A Better Title Than Pipetting"
author: "Nathan Malmberg"
date: "6/24/2019"
output:
  bookdown::html_document2:
    number_sections: FALSE
bibliography: [knitr.bib, references.bib]
csl: american-chemical-society.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Results

Put all of your data into the code block below, in the read.table function. Also,
edit the lm functions to reflect the specific pipetters you've used for your data.
When you write, don't treat the code block as a figure or table; it won't be
labeled as one, and it exists primarily so your instructor knows that you've done
the calculations correctly. You should include a reference to the worksheet and to
R in this paragraph, since you won't have a Methods section in this lab.

```{r Calculations}
pipetteframe <- read.table(header=TRUE, text="
Pipette   Volume   Before   After")
pipetteframe$Mass <- pipetteframe$After - pipetteframe$Before
biglm <- lm(Mass ~ Volume, data=pipetteframe, subset=(Pipette=="1000D"))
summary(biglm)
littlelm <- lm(Mass ~ Volume, data=pipetteframe, subset=(Pipette=="100A"))
summary(littlelm)
```

Be sure to edit the code below to properly subset your two pipetters.

```{r Plot, fig.cap="(ref:plot)"}
plot(Mass ~ Volume, data=pipetteframe, type="n",
     xlab=expression(paste("Volume (", mu, "L)")),
     ylab="Mass (g)")
points(Mass ~ Volume, data=pipetteframe, pch=21, subset=(Pipette=="1000D"))
points(Mass ~ Volume, data=pipetteframe, pch=23, subset=(Pipette=="100A"))
abline(biglm, lty=1)
abline(littlelm, lty=2)
legend(0,1,legend=c("1000D","100A"),pch=c(21,23), lty=1:2)
```

(ref:plot) Put a descriptive caption for your plot here.

Write your results here. Make sure to incorporate the concepts from the questions
in the handout, but structure the concepts in paragraph form, referring to the plot
as appropriate.

## References

```{r include=FALSE}
knitr::write_bib("base", "knitr.bib")
```


